---
title: "Homework 1"
author: "Eileen Shea"
date: "September 21, 2018"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

Here are two chunks of code that create the described data frame:

```{r var_create}
set.seed(123)
x = runif(10, min = 0, max = 5)
```

```{r df_create}
hw1_prob1_df = tibble(
  vec_numeric = x,
  vec_logical = c(FALSE, TRUE, TRUE, TRUE, TRUE, FALSE, TRUE, TRUE, TRUE, TRUE),
  vec_char = c("You", "miss", "one-hundred", "percent", "of", "the", "shots", "you", "don't", "take"),
  vec_factor = factor(c("yes", "no", "no", "yes", "no", "yes", "yes", "no", "no", "yes"))
)
```

```{r means}
mean(x)
mean(FALSE)
mean(TRUE)
mean("You")
mean("miss")
mean("yes")
mean("no")
```

Taking the mean of **numeric** and **logical** variables *works*, while taking the mean of character and factor variable does not. This is because numeric and logical variables have a number assigned to them (in the case of FALSE and TRUE this is 0 and 1 respectively), while character and factor variables do not have assigned number values to summarize in a statistic like the mean.









