---
title: "Homework 1"
author: "Eileen Shea"
date: "September 21, 2018"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

Here are two chunks of code that create the described data frame:

```{r var_create}
set.seed(123)
x = runif(10, min = 0, max = 5)
```

```{r df_create}
hw1_prob1_df = tibble(
  vec_numeric = x,
  vec_logical = c(FALSE, TRUE, TRUE, TRUE, TRUE, FALSE, TRUE, TRUE, TRUE, TRUE),
  vec_char = c("You", "miss", "one-hundred", "percent", "of", "the", "shots", "you", "don't", "take"),
  vec_factor = factor(c("yes", "no", "no", "yes", "no", "yes", "yes", "no", "no", "yes"))
)
```

Next is a code chunk that tries to take the mean of the variables in the data frame:

```{r means}
mean(hw1_prob1_df$vec_numeric)
mean(hw1_prob1_df$vec_logical)
mean(hw1_prob1_df$vec_char)
mean(hw1_prob1_df$vec_factor)
```

Taking the mean of **numeric** and **logical** variables *works*, while taking the mean of character and factor variables does not. This is because numeric and logical variables have a number assigned to them (in the case of FALSE and TRUE this is 0 and 1 respectively), while character and factor variables do not have assigned number values to summarize through a statistic like the mean.

But we can try to convert variables from one type to another, as in the below code chunk:

```{r num_convert, eval=FALSE}
as.numeric(hw1_prob1_df$vec_logical)
as.numeric(hw1_prob1_df$vec_char)
as.numeric(hw1_prob1_df$vec_factor)
```

When we apply the as.numeric function to the logical variables they are numerically represented by 0 (for FALSE) and 1 (for TRUE); when we apply the as.numeric function to the factor variables they are then numerically represented by 1 (no) and 2 (yes). However, the character variables, even when the as.numeric function is applied, are all assigned NA (not available, i.e. missing values).

Here is a second code chunk demonstrating conversion between variable type:

```{r second_convert}
as.numeric(as.factor(hw1_prob1_df$vec_char))
as.numeric(as.character(hw1_prob1_df$vec_factor))
```

When you convert a character variable to a factor variable and then a numeric variable, each character element ultimately gets assigned a unique number in the end. Alternatively, when you convert a factor variable to a character variable and then a numeric variable, each factor variable ultimately ends up being assigned an NA value.


## Problem 2

Below is the code chunk that creates the data frame for Problem 2:

```{r df_create2}
hw1_prob2_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  vec_logical = ifelse(x + y > 0, TRUE, FALSE),
  vec_numeric = as.numeric(hw1_prob2_df$vec_logical),
  vec_factor = as.factor(hw1_prob2_df$vec_logical)
)
```



